---
title: A simple model for dynamics of beliefs
author: "gpc"
date: "05/10/2014"
output: 
  html_document:
    keep_md: true
---

<style>
    iframe{
        border: none
    }
</style>

```{r setup, echo=FALSE, message=FALSE}
library(knitr)
opts_knit$set(animation.fun = hook_scianimator)
opts_chunk$set(fig.path='figures/model00-', collapse=TRUE)
library(xtable)
library(stringr)
library(Rgraphviz)
library(reshape2)
library(ggplot2)
```

 
## Society as a graph 

Consider for illustration a small society with $5$ individuals names `JP`, `AA`, `J`, `B` and `JBM`. 
It is represented by a graph such as the one below. The vertices are the individuals and the edges are the connections between individuals.

```{r, echo=FALSE}
a <- "JP"
b <- "AA"
c <- "J"
d <- "B"
e <- "JBM"
members <- c(a, b, c, d, e)
M <- matrix(
  c(1,  1,	1,	0,
    NA,	1,	1,	0,
    NA,	NA,	1,	1,
    NA, NA,	NA,	1),
  nrow=4, byrow=TRUE)
rownames(M) <- members[-length(members)]
colnames(M) <- members[-1]
```

```{r graph, fig.height=3, echo=FALSE}
library(Rgraphviz)
g <- new("graphNEL", nodes=members, edgemode="undirected") 
for(i in 1:nrow(M)){
  connections <- which(M[i,]>0)
  if(length(connections)){
    for(j in connections){
      g <- addEdge(rownames(M)[i], colnames(M)[j], g, M[i,j])
    }
  }
}
#plot(g)
g <- layoutGraph(g)
nodeRenderInfo(g) <- list(fill=c("JP"="lightblue"))
renderGraph(g)
```

This graph is easy to plot in R using the `graph` and `Rgraphviz` packages. 
One starts with a matrix of connections `M`:

```{r}
M
```

prettier rendered like this (for those who are not familiar with R outputs):

```{r, results='asis', echo=FALSE}
tab <- print(xtable(M, digits=0, align=c("r|","c|","c|","c|","c")), 
             print.results=FALSE, floating=FALSE, comment=FALSE, hline.after=0:3)
tab <- str_replace_all(tab, "tabular", "array")
cat(paste0("$$", tab, "$$"))
```

A `0` means no connection between the two individuals, a `1` means connection between the two individuals (connected individuals meet sometimes). 
Then the graph is constructed and plotted by the following code:

```{r chunk2, ref.label='graph', eval=FALSE}
```

Individual `JP` will play a special role later, that's why we use a color for his vertex. 

For fun, we can generate a Javascript graph with the `d3Network` package:

```{r, results='asis'}
library(d3Network)
from <- sapply(str_split(edgeNames(g), "~"), function(pair) pair[1])
to <- sapply(str_split(edgeNames(g), "~"), function(pair) pair[2])
d3SimpleNetwork(data.frame(from,to), file="d3graph.html", 
                fontsize=13, linkDistance=150, 
                iframe=TRUE, width=400, height=300)
```


## Belief updating

Consider a topic of interest in the society. Each individual has its own opinion about this topic, represented by a score between $0\%$ and $100\%$.  
When two connected individuals meet, their opinion is updated according to a *trusting parameter*. 
More precisely, denote by $x_i(t)$ the opinion of each individual $i$ at time $t$. Then, when individual $i$ meet individual $j$ at time $t$,  its opinion is updated according to the following rule:
$$x_i(t^+) =  (1-\alpha_{ij}) x_i(t) + \alpha_{ij} x_j(t)$$
where $\alpha_{ij} \in [0,1]$ is the trusting parameter representing the degree of individual $i$ confidence in individual $j$. 
In other words, the updated opinion $x_i(t^+)$ is the average of the two opinions $x_i(t)$ and $x_j(t)$ weighted by the trusting parameter $\alpha_{ij}$.  
It is understood that individual $j$ opinion is similarly updated:
$$x_j(t^+) =  (1-\alpha_{ji}) x_j(t) + \alpha_{ji} x_i(t)$$

<u>Extreme trusting parameters</u>:

- $\alpha_{ij}=0$ *($i$ stubborn to $j$)*: in case $\alpha_{ij}=0$, opinion of individual $i$ does not change when he meets individual $j$
- $\alpha_{ij}=1$ *($i$ dominated by $j$)*: in case $\alpha_{ij}=1$, opinion of individual $i$ becomes opinion of individual $j$ when they meet

For our illustration we use the following matrix of trusting parameters (there's no trusting parameter between unconnected individuals):

```{r, echo=FALSE}
alpha <- as.matrix(structure(list(JP = c(NA, 100L, 60L, 100L, NA), AA = c(0L, NA, 
                                                            10L, 50L, NA), J = c(0L, 90L, NA, 50L, 20L), B = c(0L, 90L, 
                                                                                                                10L, NA, 0L), JBM = c(NA, NA, 10L, 20L, NA)), .Names = members, class = "data.frame", row.names = members))/100
```

```{r}
alpha
```

```{r, results='asis', echo=FALSE}
tab <- print(xtable(alpha, digits=1, align=c("r|", "c|", "c|","c|","c|","c")), 
             print.results=FALSE, floating=FALSE, comment=FALSE, hline.after=0:4)
tab <- str_replace_all(tab, "tabular", "array")
cat(paste0("$$\\alpha=", tab, "$$"))
```

and we use the following initial values of the opinions at time $t=0$:

```{r}
individuals <- colnames(alpha)
inits <- setNames(c(100, 50, 10, 5, 15), individuals)
inits
```

Thus :

- `JP` is stubborn to everyone: his trusting parameters $\alpha_{1,j}$ are equal to $0$. That means his opinion will never change and he starts with extreme opinion $x_1(0)=100\%$. He's playing the role of a media in the society.

- `AA` is dominated by `JP`: her trusting parameter $\alpha_{2,1}$ is equal to $1$. That means she will adopte whenever she meets `JP`. She is a credulous person: her other trusting parameters are $0.9$. She starts with opinion $x_2(0)=50\%$. 

- `J` is a bit credulous about the claims of `JP` ($\alpha_{3,1}=60\%$) but not very credulous about the claims of his other friends ($\alpha_{3,j}=0.2$). He starts with opinion $x_3(0)=10\%$. 

- `B` is dominated by `JP` ($\alpha_{4,1}=100\%$), half-credulous in regards to `AA` and `J` ($\alpha_{4,j}=50\%$), and not very credulous in regards to `JBM` ($\alpha_{4,5}=20\%$). He starts with opinion $x_4(0)=5\%$. 

- `JBM` is not very credulous in regards to `J` ($\alpha_{5,3}=20\%$) but gives no credence to `B` ($\alpha_{5,4}=0\%$). He starts with opinion $x_5(0)=15\%$. 

Well, let me mention that this example is a joke inspired by an episode of my personal life. The belief of interest is "*Gilbert PC is crazy*", but it has no importance for the purpose.  

## Occurences of meetings

We simulate the occurences of meetings between individuals as simple as possible: at regular time intervals, we pick at random one connection between a pair of individuals. 

```{r, fig.width=13}
T <- 200
X <- matrix(NA, nrow=nrow(alpha), ncol=T)
rownames(X) <- rownames(alpha)
X[,1] <- inits
update <- function(X, t){
  edge <- str_split(edgeNames(g), "~")[[sample.int(numEdges(g),1)]]
  i <- edge[1] ; j <- edge[2]
  xi <- X[i,t-1] ; xj <- X[j,t-1]
  X[i,t] <- (1-alpha[i,j])*xi + alpha[i,j]*xj
  X[j,t] <- (1-alpha[j,i])*xj + alpha[j,i]*xi
  others <- setdiff(rownames(X), edge)
  X[others,t] <- X[others,t-1]
  return(X)
}
for(t in 2:T){
  X <- update(X,t)
}
sims <- data.frame(time=1:T, t(X))
ggdata <- melt(sims, id.vars="time", variable.name="individual", value.name="opinion")
ggplot(ggdata, aes(x=time, y=opinion, color=individual)) + 
   geom_line() + 
  xlim(c(1,T)) + ylim(c(0,100))
```



## A two-stubborns example

Let's modify the trusting parameters of `JBM` as well as his initial opinion value. We set all to $0$, so that `JBM` is now the subborn opposite to `JP`. 

```{r, fig.width=13}
inits["JBM"] <- 0
alpha["JBM","J"] <- 0
T <- 1000
X <- matrix(NA, nrow=nrow(alpha), ncol=T)
rownames(X) <- rownames(alpha)
X[,1] <- inits
update <- function(X, t){
  edge <- str_split(edgeNames(g), "~")[[sample.int(numEdges(g),1)]]
  i <- edge[1] ; j <- edge[2]
  xi <- X[i,t-1] ; xj <- X[j,t-1]
  X[i,t] <- (1-alpha[i,j])*xi + alpha[i,j]*xj
  X[j,t] <- (1-alpha[j,i])*xj + alpha[j,i]*xi
  others <- setdiff(rownames(X), edge)
  X[others,t] <- X[others,t-1]
  return(X)
}
for(t in 2:T){
  X <- update(X,t)
}
sims <- data.frame(time=1:T, t(X))
ggdata <- melt(sims, id.vars="time", variable.name="individual", value.name="opinion")
ggplot(ggdata, aes(x=time, y=opinion, color=individual)) + 
   geom_line() + 
  xlim(c(1,T)) + ylim(c(0,100))
```




```{r}
#knit_exit()
```

You have to embed the jQuery and SciAnimator libraries in the HTML header
(between `<head>` and `</head>`), even though some browsers can tolerate it
if you insert a style sheet after `<body>`.

<link rel="stylesheet" href="https://raw.github.com/brentertz/scianimator/master/assets/css/scianimator.css">
<script src="https://ajax.googleapis.com/ajax/libs/jquery/1.7.1/jquery.min.js"></script>
<script src="http://vis.supstat.com/assets/themes/dinky/js/jquery.scianimator.min.js"></script>

Here comes a test of the Brownian motion:

```{r test, fig.show='animate', interval=.1}
for(t in 1:20){
  gg <- ggplot(subset(ggdata, time<=t), aes(x=time, y=opinion, color=individual)) + 
     geom_line() + 
    xlim(c(1,T)) + ylim(c(0,100))
  print(gg)
  }
```


```{r}
knit_exit()
```

